---
title: "p8105_hw1_lz3014"
author: "Liqi Zhou"
date: 2024-09-20
output: github_document
---

```{r}
library(palmerpenguins)
library(tidyverse)
```

# Problem 1
#### 1. load and view the `penguins` dataset:
```{r}
data("penguins", package = "palmerpenguins")
```


#### 2. description of the `penguins` dataset
- **names of variables:** `r colnames(penguins)`  
- **size of tibble:** this tibble has `r nrow(penguins)` rows and `r ncol(penguins)` columns  
- **mean flipper length:** `r mean(pull(penguins, flipper_length_mm), na.rm = TRUE)` mm  


#### 3. draw and save scatterplot of flipper_length_mm vs bill_length_mm, colored by species
```{r}
penguins |>
  drop_na(bill_length_mm, flipper_length_mm) |>
  ggplot(aes(x = bill_length_mm, 
             y = flipper_length_mm, 
             color = species)) + 
  geom_point() + 
  labs(title = "Scatterplot of flipper length vs bill length", 
       x = "flipper length (mm)", 
       y = "bill length (mm)")

ggsave("Scatterplot of flipper length vs bill length.png")
```


# Problem 2
#### 1. create a data frame (seed:0920):
```{r}
data_p2 <- tibble(
  vec_num = rnorm(10),
  vec_logi = vec_num > 0,
  vec_char = c("a","b","c","d","e","f","g","h","i","j"),
  vec_factor = factor(rep(c("level_1","level_2","level_3"), length.out = 10))
)
data_p2
```